Our next task is to provide a mapping that can translate
choreographies into the PRISM language. This section addresses
projection formally.
% The goal of this section is to address the theoretical aspect of
% this.

\mypar{Mapping Choreographies to PRISM.}  The operation of generating
endpoint code from a choreography is known as {\em
  projection}. 
% Normally, projection is defined separately for each module appearing
% in the choreography program.
We observe that to simulate a choreography interaction in
PRISM, we need to use labels on which each involved module can
synchronise. Therefore, without loss of generality, we make a slight
abuse of notation and assume that each interaction in a choreography
is annotated with a unique label (which will be used by the
projection). We call such a choreography an {\em annotated
  choreography}. For example, the choreography
$\interactBase{p}{q}\ \lambda_1: \CEnd \ +\ \lambda_2: \ifTE
  {E}{p}{\CEnd}{\CEnd}$ is going to be annotated as
  $\interactBasel{p}{a_1}{q}\ \lambda_1: \CEnd \ +\ \lambda_2: \ifTEl
    {E}{a_2}{p}{\CEnd}{\CEnd}$, with $a_1$ and $a_2$ distinct.
% %
% \MC{We need to say that we must run some standard static checks
% because, since there is branching, some terms may not be
% projectable.}
% %
We now separate the definition of projection based on whether we
are dealing with rates or probabilities.  We start with choreographies 
with rates: 
% can then define the projection of a choreography with rates into the
% PRISM language.
% 
\begin{definition}[Projection, CTMC]\label{def:projCTMC} Given an
  annotated choreography with rates $C$, a module $\role p$, and a
  natural number $\iota$, we define the function $\proj$ as:
  \begin{displaymath}\small
        \begin{array}{lr}

          \proj (\role q,\interactl{p}{a}{\role p_1,\ldots,\role p_n}, \iota)= 
          &  \boxed{\text{if }\role q=\role p}\\[2mm]
          \qquad
          % \{ 
          \Big\{\commandBase {a_j} {\ s_{q}\!=\! \iota} {\lambda_j:\ s_{q}\!=\!
          s_{q} +1+\sum_{k=1}^{j-1}\mathsf{nodes}({C_k})}\ \&\ \projE
          {E_j}q\Big\}_{j\in J}
          \\
          \qquad\cup\ \bigcup_{j} \proj (\role q, C_j, \iota+1+\sum_{k=1}^{j-1}\mathsf{nodes}({C_k}))
          \\\\

          \proj (\role q,\interactl{p}{a}{\role p_1,\ldots,\role p_n}, \iota)= 
          &  \!\!\!\!\!\! \!\!\!\!\!\!\boxed{\text{if }\role q\in\{\role p_1,\ldots,\role p_n\}}\\[2mm]
          \qquad
          % \{ 
          \Big\{\commandBase {a_j} {\ s_{q}\!=\! \iota} {1:\ s_{q}\!=\!
          s_{q} +1+\sum_{k=1}^{j-1}\mathsf{nodes}({C_k})}\ \&\ \projE
          {E_j}q\Big\}_{j\in J}
          \\
          \qquad\cup\ \bigcup_{j} \proj (\role q, C_j, \iota+1+\sum_{k=1}^{j-1}\mathsf{nodes}({C_k}))
          \\\\

          \proj (\role q,\interactl{p}a{\role p_1,\ldots,\role p_n}, \iota)=\ \proj (\role{q}, C_1, \iota)
          &  \hspace{-1.8cm}\boxed{\text{if }\role q\not\in\{\role p, 
            \role p_1,\ldots,\role p_n\}}
          % \\[2mm]
          % \qquad\ \proj (\role{q}, C_1, s)=
          % \ \cup\
          % \proj (\role{q}, C_2, s+\mathsf{nodes}(C_1)) 
          \\[2mm]
          \qquad\text{such that } \forall i,j.\ \proj (\role{q}, C_i, \iota)= \proj (\role{q}, C_j, \iota)
          \\\\

          \proj (\role q,\ifTE {E}{p}{C_1}{C_2}, \iota) = 
          &  \boxed{\text{if }\role q=\role p}\\[2mm]
          \qquad\left\{ 
          \begin{array}{lll}
            \commandBase {} {s_{q}\!=\! \iota\ \&\ E}{\ 1: s'_{q}\!=\! \iota+1},\\ 
            \commandBase {} {s_{q}\!=\! \iota\ \&\ \mathsf{not}(E)}
            {\ 1: s'_{q}\!=\! \iota+\mathsf{nodes}(C_1)+1}
          \end{array}
          \right\}
          \\
          \qquad\cup\quad \proj (\role{p}, C_1, \iota+1)
          \quad\cup\quad
          \proj (\role{p}, C_2, \iota+\mathsf{nodes}(C_1)+1)
          \\\\

          \proj (\role q,\ifTE {E}{p}{C_1}{C_2}, \iota) = 
          &  \boxed{\text{if }\role q\neq\role p}\\[2mm]
          \qquad\proj (\role{q}, C_1, \iota+1)
          \quad\cup\quad
          \proj (\role{q}, C_2, \iota+\mathsf{nodes}(C_1)+1)

          \\\\

          \proj (\role q,\CEnd, \iota) = \emptyset

        \end{array}
      \end{displaymath}
    \end{definition}
    We go through the various cases of the definition above. The first
    three cases handle the projection of an interaction. If we project
    the first module $\role p$, then we create one command per branch,
    assigning the corresponding rate. Note that this is possible since
    we are dealing with rates. The additional variable $s_{\role q}$
    is the counter for this module and we identify uniquely this
    interaction. In order to do it consistently for follow up
    statements of the subterm choreographies, we use the function
    $\mathsf{nodes}(C)$ which returns the number of nodes of $C$,
    i.e., the number of steps of the projection function. Obviously,
    when projecting the next branch we need to consider all
    other possible branches that may have been already
    projected. Intuitively, a label and an integer (ranged by $\iota$)
    identify a node in the abstract syntax tree of a
    choreography. Also, note that we assume that from a label $a$ we
    can generate distinct sublabels $a_j$ just by adding some index
    $j$. For the sake of space, we do not define the function
    precisely, but we observe that this could also be easily defined
    via the label annotations. The second case defines the projection
    of an interaction when we are projecting one of the modules
    $\{\role p_1,\ldots,\role p_n\}$. Similarly to the previous case,
    we define a command for each branch of the interaction. However,
    the rate of each command is going to be $1$. The if-then-else
    construct is only interesting for module $\role p$: in this case,
    the module makes an internal choice based on the evaluation of the
    guard $E$. The other modules will be projected according to
    choreographies $C_1$ and $C_2$.

    As hinted above, the projection in Definition~\ref{def:projCTMC}
    would be incorrect if instead of using rates we used
    probabilities. This is simply because we cannot force both
    $\role p$ and $\{\role p_1,\ldots,\role p_n\}$ to take the same
    branch with the probability distribution of the $\lambda_i$'s. To fix this problem, we have the following definition
    instead:
    % 
    \begin{definition}[Projection, DTMC]\label{def:projDTMC} Given an annotated
      choreography with probabilities $C$, a module $\role p$, and a
      natural number $\iota$, we define $\proj$ as:
      \begin{displaymath}\small
        \begin{array}{lr}

          \proj (\role q,\interact{p}{\role p_1,\ldots,\role p_n}, \iota)= 
          &  \boxed{\text{if }\role q=\role p}\\[2mm]
          \qquad
          \left\{
          \begin{array}{lll}
            \commandBase {} {s_{q}\!=\! \iota}{\ \sum_{j\in J} \lambda_j: s'_{q}\!=\! \iota+1+j},\\ 
            \{\commandBase {l_j} {s_{q}\!=\! \iota+1+j}
            \ 1: s'_{q}\!=\! \iota+1+\sum_{k=1}^{j-1}\mathsf{nodes}({C_k})\ \&\ \projE
            {E_j}q\}_{j\in J}
          \end{array}
          \right\}
          \\
          \qquad\cup\ \bigcup_{j} \proj (\role q, C_j, s+1+\sum_{k=1}^{j-1}\mathsf{nodes}({C_k}))
          \\\\

        \end{array}
      \end{displaymath}
      The other cases of the definition are equivalent to those in
      Definition \ref{def:projCTMC}.
    \end{definition}
    The fix is immediate: module $\role p$ takes a (probabilistic)
    internal decision on the $j^{\text{th}}$ branch and then
    synchronises on label $l_j$ with $\{\role p_1,\ldots,\role p_n\}$.       

\smallskip

\mypar{Correctness.} We now show that our projection operations are
correct with respect to the semantics of choreographies and PRISM. In
the sequel, $S_+$ is a state $S$ extended with the extra variables
$s_{\role q}$ (for each module in a choreography) used by the
projection. In the projection, we utilize alphabetized parallel
composition $\pp$, wherein modules synchronize solely on labels that
appear in both modules.
% Moreover, we say that a choreography is {\em well-formed} if the
% behaviour in each branching for modules who do not participate to
% the interaction is identical.
\begin{theorem}[EPP]\label{thm:epp}
  Given a well-formed annotated choreography $C$, we have that
  $(S,C) \red{\lambda} (S', C')$ iff
  $\pp_{\role q}\ \proj (\role q, C,\iota)\vdash S_{+}
  \red{\lambda} S_{+}'$.
\end{theorem}
\begin{proof} The proof must be separated into whether we deal with
  rates and with probabilities. Let us focus on the projection in
  Definition~\ref{def:projCTMC}. The proof proceeds by induction on
  the term $C$. For $C=\interact{p}{\role p_1,\ldots,\role p_n}$, the
  key case, we clearly have that, for any state $S$ there exists $S'$
  such that $(S,C)\red{\lambda_j}(S',C_j)$. We need to show two
  things: first, that the projection
  $\pp_\role q\ \proj{(\role{q},C,\iota)}$ of $C$ can make the same
  transition; second, that if the projection makes a transition, it
  must be corresponding to that of the choreography above.

  We now observe that this state of the CTMC in the translation is
  uniquely identified by the counter $\iota$. 
  % Technically, this follows by the definition of projection and
  % annotation.
  Moreover, the uniqueness of the label $a$ makes sure that all and
  only those modules involved in this interaction synchronise with
  this action (this shows from the rules in
  Figure~\ref{fig:semantics}).  As a consequence, the commands
  generated by this step of the translation are such that any state
  $S_+$ is exclusively going to enable these commands (because of the
  guard $s_{\role q}=\iota$) which obviously implies that it must be
  done with rate $\lambda_j$ applying the rules in
  Figure~\ref{fig:semantics}. The same argument can be applied for the
  opposite direction. 
  % Note that having a unique label for each step of the choreography
  % is only necessary for recursion. Without recursion, having a
  % unique state identifying this particular step of the choreography
  % would have been sufficient.
  The other cases are similar. The case for DTMC is also similar.



  %       \newpage We prove each direction separately.
  %       \begin{itemize}
  %       \item (only if). Assume that
  %         % 
  %   $$(S, \interact{p}{\role p_1,\ldots,\role
  %     p_n})\red{\lambda_j}(S[\sigma(E_j)/x_j], C_j)$$
  %   % 
  %   and let us consider the projection of the term
  %   %
  %   $$\interact{p}{\role p_1,\ldots,\role p_n}$$
  %   % 
  %   Given some fresh $l_1,\ldots, l_n$, we generate the following
  %   commands for each $q$ in
  %   $\{\role p, \role p_1,\ldots,\role p_n\}$:
  %   % 
  %   $$
  %   \Big\{\commandBase {l_j} {\ s_{q}\!=\! s} {\kappa_j:\ s_{q}\!=\!
  %     s_{q} +1+\sum_{k=1}^{j-1}\mathsf{nodes}({C_k})}\ \&\ \projE
  %   {E_j}q\Big\}_{j\in J}
  %   $$
  %   %
  %   where $\kappa_j=\lambda_j$ if $q\neq\role p$ and $\kappa_j=1$
  %   otherwise. 
  %   % 
  %   Because the labels $l_j$ are fresh and the state counter is unique
  %   to this interaction, these are the only commands that can
  %   synchronise together: this can be shown from the rules defining
  %   the semantics of PRISM. Additionally, since all rates are set to 1
  %   besides the commands generated for role $\role p$, the transition
  %   will also be labelled with $\lambda_j$.


  % \item (if). In the opposite direction, we have that 
  %   % 
  %   $$\proj (*, \interact{p}{\role p_1,\ldots,\role p_n})\vdash
  %   S\uplus S_{+}\red{\lambda} S'\uplus S_{+}'$$
  %   %
  %   Again, given some fresh $l_1,\ldots, l_n$, the projection that
  %   reduces must be such that for each $q$ in
  %   $\{\role p, \role p_1,\ldots,\role p_n\}$:
  %   % 
  %   $$
  %   \Big\{\commandBase {l_j} {\ s_{q}\!=\! s} {\kappa_j:\ s_{q}\!=\!
  %     s_{q} +1+\sum_{k=1}^{j-1}\mathsf{nodes}({C_k})}\ \&\ \projE
  %   {E_j}q\Big\}_{j\in J}
  %   $$
  %   %
  %   where $\kappa_j=\lambda_j$ if $q\neq\role p$ and $\kappa_j=1$
  %   otherwise. 
  %   % 
  %   Again, the freshness of the labels $l_j$ together with the working
  %   states $s_qq$

  %   Because the labels $l_j$ are fresh and the state counter is unique
  %   to this interaction, these are the only commands that can
  %   synchronise together: this can be shown from the rules defining
  %   the semantics of PRISM. Additionally, since all rates are set to 1
  %   besides the commands generated for role $\role p$, the transition
  %   will also be labelled with $\lambda_j$.

  % \end{itemize}
\qed
\end{proof}

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:
