\documentclass[11pt]{article}

% Page layout
\usepackage{geometry}
\geometry{margin=1in}

% Formatting
\usepackage{xcolor}
\usepackage{setspace}
\usepackage{enumitem}
\usepackage{amssymb}
\usepackage{url}

% Boxes for reviewer comments
\usepackage{tcolorbox}
\tcbuselibrary{skins,breakable}

% Markers
\newcommand{\addressed}{\textcolor{green!60!black}{\checkmark}\ }
\newcommand{\notaddressed}{\textcolor{red!70!black}{\texttimes}\ }

\newcommand{\defin}{\mathcal D}

\newcommand{\sem}[1]{\{\![#1]\!\}}
\newcommand{\eval}[2]{#1\!\!\downarrow_{#2}}

\newcommand{\mypar}[1]{\noindent{\bf #1}}

\newcommand{\state}{\mathsf{state}}

\newcommand{\proj}{\mathsf{proj}}
\newcommand{\projE}[2]{#1\downarrow_{#2}}

\newcommand{\sconnected}{\mathsf{sConn}}
\newcommand{\hmodules}{\mathsf{hMods}}


\newcommand{\topannotation}[2]{\mathsf{top}_{#1}(#2)}

\newcommand{\marco}[1]{\textcolor{cyan}{{\footnotesize Marco: #1}}}
\newcommand{\adele}[1]{\textcolor{blue}{{\footnotesize Adele: #1}}}

\definecolor{code_indent}{HTML}{CCCCCC}

\newcommand{\indentrule}{\color{code_indent}\vrule\hspace{2pt}}

% Syntax
\definecolor{bloodyred}{rgb}{0.6,0.1,0.1}
\newcommand{\role}[1]{\textcolor{bloodyred}{\mathtt{#1}}}
%\newcommand{\role}[1]{\mathtt{#1}}
\newcommand{\nr}{\role p}
\newcommand{\nrr}{\role q}
\newcommand{\nrrr}{\role r}

\newcommand{\Var}{\mathsf{Var}}
\newcommand{\Val}{\mathsf{Val}}

\newcommand{\interactBase}[2]{{#1}\rightarrow \{{#2}\}}
\newcommand{\interactBasel}[3]{#1\rightarrow^{#2} \{#3\}}
\newcommand{\interact}[2]{\interactBase{#1}{#2}:\,\Sigma_{j\in J}\lambda_j: u_j;\ C_j}
\newcommand{\interactl}[3]{\interactBasel{#1}{#2}{#3}:\,\Sigma_{j\in J}\lambda_j: u_j;\ C_j}
\newcommand{\doubleand}{\text{\&\&}}
\newcommand{\allsynchName}{\textcolor{bloodyred}{\mathtt{allsynch}}}
\newcommand{\allsynch}[4]{\allsynchName\, \{ {\role{#1}}_{#3}:#2_{#3}\}_{#3\in #4}}

\newcommand{\command}[4] {\commandBase {#1} {#2} \Sigma_{i\in I}#3_i: #4_i}
\newcommand{\commandBase}[3] {[#1]\ #2\ \rightarrow #3}
\newcommand{\chorcommandBase}[2] {#1\ \rightarrow #2}
\newcommand{\chorcommand}[3] {\chorcommandBase {#1} \Sigma_{i\in I}#2_i: #3_i}
\newcommand{\ifTE}[4]{\textcolor{sh_keyword}{\mathtt{if}}\ #1@{#2}\ \textcolor{sh_keyword}{\mathtt{then}}\ {#3} \ \textcolor{sh_keyword}{\mathtt{else}}\ {#4}}
\newcommand{\ifTEl}[5]{\textcolor{sh_keyword}{\mathtt{if}}\ #1@^{#2}{#3} \ \textcolor{sh_keyword}{\mathtt{then}}\ {#4} \ \textcolor{sh_keyword}{\mathtt{else}}\ {#5}}
\newcommand{\codeprism}[1]{\lstinline[style=prism-color,basicstyle=\small\sffamily]{#1}}
\newcommand{\codechor}[1]{\lstinline[style=chor-color,basicstyle=\small\ttfamily,mathescape=true]{#1}}



\newcommand{\CEnd}{\mathbf{0}}
\newcommand{\pp}{|\!|}
\newcommand{\ppp}[1]{|[#1]|}
\newcommand{\cmd}[6]{[#1] #2 \rightarrow \{#3: #4 = #5\}_{#6}}

\newcommand{\defrec}{\stackrel{\textcolor{sh_keyword}{\mathsf{def}}}{=}}

  
% Semantics
\newcommand{\red}[1]{\ \longrightarrow_{#1}\ } 
\newcommand{\prismred}[1]{\ \stackrel{#1}{\rightsquigarrow}\ } 
\newcommand{\lab}[3]{\ [#1][#2:\;#3]\ } 

\definecolor{pblue}{rgb}{0.13,0.13,1}
\definecolor{pgreen}{rgb}{0,0.5,0}
\definecolor{pred}{rgb}{0.9,0,0}
\definecolor{pgrey}{rgb}{0.46,0.45,0.48}
\definecolor{keywordColor}{rgb}{0.50,0.00,0.33}
\definecolor{stringColor}{rgb}{0.16,0.00,1.00}
\definecolor{lineNumberColor}{rgb}{0.47,0.47,0.47}

\definecolor{carminepink}{rgb}{0.92, 0.3, 0.26}
\definecolor{jesuscolour}{rgb}{0.3, 0.92, 0.26}
\newcommand{\AV}[1]{\todo[inline, color=carminepink]{\textbf{AV}:~#1}}
\newcommand{\MC}[1]{\todo[inline, color=jesuscolour]{\textbf{MC}:~#1}}


% Very light grey background
\definecolor{reviewgray}{gray}{0.95}

\newtcolorbox{reviewercomment}{
  breakable,
  colback=reviewgray,
  colframe=black!20,
  boxrule=0.3pt,
  arc=2pt,
  left=6pt,
  right=6pt,
  top=6pt,
  bottom=6pt
}

% Commands
\newcommand{\revcomment}[1]{%
  \begin{reviewercomment}
  \textbf{Reviewer comment.} #1
  \end{reviewercomment}
}

\newcommand{\revresponse}[2][]{%
  \par\noindent\textbf{Response.} #1 #2
}

\newcommand{\revchanges}[2][]{%
  \par\noindent\textbf{Changes in the manuscript.} #1
  #2\par\bigskip
}

\begin{document}

\title{Response to Reviewers}
\author{A PROBABILISTIC CHOREOGRAPHY LANGUAGE FOR PRISM}
\date{}
\maketitle
\vspace{-0.5\baselineskip}
\begin{center}
{\large Marco Carbone and Adele Veschetti}
\end{center}

\section*{General Response}

We thank the Editor and the Reviewers for their careful reading of the manuscript and
for their detailed and constructive feedback. We have revised the paper accordingly
and believe that the changes substantially improve both clarity and technical depth.

One major change in this revision concerns the definition of
projection. In the previous version, the projection relied on explicit
counters (denoted by $\iota$) to generate fresh identifiers; this
mechanism was cumbersome and could be error-prone in proofs. We now
adopt a more direct approach by annotating choreographies with unique
labels, which the projection can use directly. This yields a cleaner
and more easily readable projection in the theoretical
development. For practical reasons, our implementation still uses
counters; these can be seen as a concrete realisation of the
label-generation mechanism that the paper now abstracts into the
choreography annotations.

Below we address each reviewer point by point. Minor comments have
been omitted from the reply, but addressed.


\newpage

% ============================================================
\section*{Reviewer 1}


\revcomment{
I found the technical result establishing operational equivalence between choreographies
and their projections rather weak. [...] This may ultimately be irrelevant, but I believe the issue should be discussed in the paper. I would also expect a result establishing
operational correspondence for a sequence of reductions.
}

\revresponse{ We thank the reviewer for highlighting the mismatch
  between standard operational-correspondence results for
  choreographies and this paper focussing on PRISM.  PRISM modules are
  declarative (a fixed set of commands whose enabling depends on the
  current state), whereas choreographies have an explicit control-flow
  syntax; therefore, in the projection, control-flow is reified into
  the (extended) runtime state via the distributed program-counter
  variables $s_{\nr}$, yielding an extended state space $S_+$. As a
  consequence, reductions of the choreography correspond to
  transitions on $S_+$ in the projected model rather than to syntactic
  evolution of PRISM modules.
%
In the revised manuscript we make this design choice explicit and discuss the resulting asymmetry (projection-after-reduction may yield fewer commands than reduction-after-projection). At the end of Section 5, we include a concrete example illustrating that, while the set of commands may differ, the projected models exhibit the same observable behaviour up to a suitable PRISM-induced equivalence; we leave the identification of the most appropriate equivalence (e.g., probabilistic bisimulation) and the corresponding strengthening of the theorem to future work.
}

\revchanges{ We expanded the discussion around Theorem 1 to clarify
  (i) why standard reduction-to-reduction correspondence does not
  directly apply to PRISM, (ii) how control-flow information is
  encoded in the extended state $S_+$ via the distributed
  program-counter variables, and (iii) why the correspondence
  statement is formulated in terms of transitions on $S_+$ rather than
  syntactic changes of modules. We also added a detailed remark and
  worked example to make the asymmetry and its implications explicit,
  and we added a paragraph outlining how a future extension based on
  probabilistic bisimulation could capture correspondence across
  different command sets.  }

\revcomment{ Furthermore, the proof of this correspondence (Theorem 1)
  reads more like a proof sketch than a complete proof; a more
  detailed and carefully paced proof would be preferable.  }

\revresponse{ We thank the reviewer for this suggestion. In the
  revised version, we have substantially expanded the proof of Theorem
  1 and made its structure more explicit, in particular by spelling
  out the interaction case in detail and by making clear the key
  invariants on the extended state $S_+$ used in the correspondence
  argument. To keep the exposition readable, we do not enumerate every
  routine PRISM semantic step, but we provide enough detail for each
  case to be checked unambiguously.  }

\revchanges{ Proof of Theorem 1 rewritten and expanded; interaction
  case detailed; key invariants on $S_+$ made explicit.  }



\revcomment{
I have also some concerns about the way in which the choreography state is projected on
different participants.
}

\revchanges{ Definitions~4 (CTMC) and~5 (DTMC) have been updated.  }

% ------------------------------------------------------------

\revcomment{
When giving examples, the use of prime variables is not clear at this point.
}

\revchanges{ We added a brief clarification in the manuscript.  }


\revcomment{
“this is the first probabilistic choreography language that is not a type abstraction”.
What does this mean?
}

\revchanges{ Comment removed.  }


\revcomment{
I did not understand the explanation of foreach. What is the role of k? What is op?
A formal encoding would be welcome.
}

\revresponse{
  We apologise for the limited explanation of this construct. The basic idea is to provide an operator that quantifies over indices via the predicate $op(i)$. The formal translation is straightforward: we add one update for each role, assuming that the relevant roles can be determined statically.
}

\revchanges{ We have updated the text where the foreach construct is
  discussed, with a more precise explanation of what it does.  }


\revcomment{ I am puzzled about the notion of Non-deterministic
  Synchronization. [...] I have a couple of questions here:

  \begin{enumerate}
  \item Does the specification determine the direction of
    communication? In the provided example, it seems that the
    interaction could also be expressed in the opposite direction,
    i.e., $q \rightarrow {p}$. Is the direction of the interaction
    relevant, or is it never relevant?
  \item Does any synchronization specification yield a correct
    choreography? Should there not be a requirement for some
    well-formedness condition on the allsync constructor?
  \item Could you provide a formal definition of the encoding?
  \end{enumerate}
}

\revresponse{The terminology we used is incorrect, thanks for pointing
  this out. . The allsync is not "nondeterministic" but it is still a
  probabilistic choice. We have rewritten (and expanded) the entire
  paragraph with a clearer explanation. As for the detailed questions:
  \begin{enumerate}
    
  \item in our choreography language, writing $q\longrightarrow\{p\}$
    or $p\longrightarrow\{q\}$ is the same. However, it has an impact
    on the projection, since we need to distinguish one of the modules
    (the one on the left) because it has a different projection.
    
  \item The notion of strong connectedness can be extended to
    $\allsynchName$, as it is, after all, an interaction between the
    modules $\nr_i$. However, this is not necessary if we first
    translate it into standard interactions (and if-then-else
    constructs), since strong connectedness can then be checked at
    that level. In the new example we have provided, strong
    connectedness does not hold because the nested if-then-else
    constructs are not connected. As this part is not a formal
    development but rather a section justifying how our implementation
    extends the theoretical framework, we do not consider it necessary
    to present a connected choreography here. In any case (as noted in
    a comment added to the projection section), these choreographies
    can be transformed into connected ones by introducing additional
    interactions.

    
  \item the translation should be straightforward once there is a
    precisely defined language for guards. See new text added. 
 \end{enumerate}
}

\revchanges{ The paragraph has been rewritten.  }


\revcomment{
Example 3: The modules in this example share variables (y and x). I wonder whether there is a choreography that describes their behavior, since my understanding is that variables are local to modules in choreographies
}

\revresponse{
In the example, $x$ and $y$ are indeed owned by two different modules and are therefore shared only in read mode. In PRISM, all modules can read all variables, but only the owning module can modify its own variable. This is consistent with our choreographic reading, where updates are only performed as part of an interaction, so the write is still attributed to a specific role in that interaction. The only caveat is that, in choreographies, guards in conditionals must be local (i.e., evaluated by a single role). Thus, for this example, we could simulate the interactive execution (sync on $a$) by nesting two if-then-else constructs. However, since choreographies focus only on interactions, the internal actions (no label) cannot be simulated.
}

\revcomment{
The choice of annotating choreographies with labels seems somewhat ad hoc. Why are they not introduced as a syntactic element [...] given that they are necessary and play a role in the implementation?
}

\revresponse{
It is certainly true that labels are a key aspect of the implementation. However, the purpose of this work is to define a higher-level language that abstracts away from such implementation details. Therefore, labels are part of the projection algorithm and are irrelevant to the programmer who writes a choreography.
}


\revcomment{
Projection of conditionals: I do not understand why the conditional is not projected.
}

\revresponse{
We believe this is a misunderstanding, and we are grateful for the opportunity to clarify it. The conditional \emph{is} projected, but only in the module of the evaluating role $\mathsf p$. This is consistent with the intended semantics: the guard $E$ is evaluated by $\mathsf p$ and the result deterministically selects the branch. The point that may be confusing is that we do \emph{not} require $E$ to be local to $\mathsf p$. As stated on page 5, $E$ may mention variables owned by other modules; in PRISM, all modules can read all variables, while only the owning module can modify them. Therefore, the projection places the conditional in $\mathsf p$'s module, and the guard can still refer to remote variables, exactly as in the PRISM syntax.
}

\revcomment{
Projection of choreography variable X: This relies on defs(X). [...]
}

\revresponse{
We agree with the reviewer. }

\revchanges{
In the revised formulation, counters are replaced by
explicit labels on choreographies, which makes the treatment of variables cleaner and more uniform. We have also expanded the proof for this case and
included more details to make the reasoning fully explicit.
}


\revcomment{
Projection of updates: Are you assuming that updates at the choreography level are well-formed in some way? For instance, could an assignment y' = x occur where y belongs to p and x to q? If not, this restriction should be stated clearly. Moreover, from which element of the choreography is the assignment derived? If I have missed this, it should nonetheless be clarified explicitly in the paper.
}

\revresponse{
As discussed above, assignments such as $y' = x$ are permitted even when $x$ is located at $\nr$ and $y$ is located at $\nrr$. At the choreography level, the restriction is simply that, in an interaction $\interact{\nr}{\nr_1,\ldots,\nr_n}$, every update in each $u_j$ assigns to a variable owned by one of $\nr, \nr_1,\ldots,\nr_n$.
}
\revchanges{
We have added a clear assumption on page 6 (new PDF).
}

\revcomment{
Definition 4: Is it not the case that the choreography is annotated?
}

\revresponse{
Interactions are now also annotaed. See new version. 
}


\revcomment{
  - How are variables quantified? Are $S$, $S'$, and $\iota$ universally quantified?
}

\revresponse{
  Yes, although now $\iota$ has disappeared.
}

\revcomment{ The correspondence theorem (Theorem 1) appears somewhat
  weak [...]  }

\revresponse{ See our reply about your first comment.  }

\revcomment{
Although I can follow the reasoning, I wonder what the fundamental limitation is here.
}

\revchanges{
Added explanation on page 24.
}

\newpage

% ============================================================
\section*{Reviewer 2}


\revcomment{
The main changes as stated by the authors at the end of the Introduction are:
[...] The paper does not help the reader in identifying them.
}

\revresponse{ Here is a list of the major differences between this
  latest version and the conference version:
  \begin{itemize}
  \item presentation of three different macros that are used in the
    implementation
  \item a new, cleaner, projection function that no longer uses
    counters for the control state
  \item detailed proof of Theorem 1
  \item new case study: Simple Peer-To-Peer Protocol
  \item new case study: Synchronous Leader Election
  \item new case study: Dining Cryptographers
  \item new case study: Code Size Comparison
  \item more detailed discussions and related work
  \end{itemize}
}


\revcomment{ "The introduction should be more clear about which kind
  of concurrent process is amenable for choreographies, and for which
  class of concurrent processes choreographies are not suitable."  The
  comment still applies for the new introduction.  }

\revchanges{We have added a paragraph in the introduction. }

  
\revcomment{
The experimental analysis in Section 7 needs to be improved.
}

\revchanges{
  This part has been updated accordingly. 
}

\revcomment{
Section 6 is irrelevant and can be removed.
}

\revchanges{
  Removed. 
}


\revcomment{
It is confusing that you use lambda1,2 as label in page 2 [...]
}
%
\revchanges{
 Fixed by rewriting the whole example paragraph in the introduction. 
}

\revcomment{ [...]
the text uses sometimes the math symbols and sometimes the full name of the greek letters lambda, mu, theta. [...]
}
%
\revchanges{
 Fixed.
}

\revcomment{
  I don't know if this intentional, but if one wants to specify no update, one has to write x' = x (a void update). In PRISM one can simply write true, if I recall properly. This is related also to the syntax of PRISM introduced in 4.1.
}

\revresponse{
  From a theoretical point of view, there is no difference, therefore we will leave this as it is. Our implementation allows also True. 
}

\revcomment{
  Are you sure Def. 1 works as intended? What happens if x appears in u? For example, [x' = 0, y' = x]. In Prism, y will be updated to the value of x in S (see remark at the bottom of \url{https://www.prismmodelchecker.org/manual/ThePRISMLanguage/Commands}), but in your case it will be set to the new value of x (0). This is also relevant to the definion of mu in page 9 and needs to be fixed. 
}

\revresponse{
 Yes sorry, our definition is incorrect. 
}

\revchanges{
 We have now updated the definition so that it handles multiple updates at the same time. 
}

\revcomment{
General remark: the way you define unique identifiers for the control points in the choreographies is not self-explainable. Readers familiar with bulding control flow graphs and similar graphical structures from code will get it, but other readers may find it cryptic. I would suggest enriching the explaination of how control points sq are correctly created. }

\revchanges{
 Added some text. 
}


\revcomment{
  "he projected update uj ↓q acts as a filter on the list of updates in uj" is this defined in the paper?
}


\revchanges{
  We have added further comments on this. 
}


% (Continue similarly for remaining detailed comments as needed.)

\newpage

% ============================================================
\section*{Reviewer 3}


\revcomment{
The pros and cons of choreographic programming as opposed to process-oriented
programming are not elaborated upon very much.
}

\revchanges{
Added discussion on pages 2 and 25--26.
}




\revcomment{
How does this relate to the notions of implementability and realisability [...]?
}

\revchanges{ We have extended related work on page 25.  }

\revcomment{ Could the authors imagine a more thorough comparison,
  e.g., bisimilarity?  }

\revchanges{ We have added a discussion on page 17.  }

\end{document}
