Programming distributed systems
%
presents significant challenges due to their inherent complexity, and
the possibility of obscure edge cases arising from their complex
interactions.
%
Unlike monolithic systems, distributed programs involve multiple nodes
operating concurrently and communicating over networks, introducing a
multitude of potential failure scenarios and nondeterministic
behaviours.
%
One of the primary challenges in understanding distributed systems
lies in the fact that the interactions between multiple components can
diverge from the sum of their individual behaviours. This emergent
behaviour often results from subtle interactions between nodes, making
it difficult to predict and reason about a system's overall behaviour.

PRISM \cite{PRISMdoc} is a probabilistic model checker that offers a
specialised language for the specification and verification of
probabilistic concurrent systems. PRISM has been used in various
fields, including multimedia protocols \cite{multimedia}, randomised
distributed algorithms \cite{distr1,distr2}, security protocols
\cite{security1,security2}, and biological systems \cite{bio1,bio2}.
At its core, PRISM provides a declarative language with a set of
constructs for describing probabilistic behaviours and properties
within a system.
%
Given a distributed system, we can use PRISM to model the behaviour of
each of its nodes,
% For example, in a distributed voting system where
% multiple nodes collaborate to reach a consensus, we could use PRISM to
% model the behaviour of each voting node individually, % specifying
% states, transitions, and communications between nodes,
and then verify desired properties for the entire system. However,
this approach can become difficult to manage as the number of nodes
increases.

%
%Furthermore, the inherent asynchrony of distributed systems
%complicates the task of programming them as intended. Conventional
%programming paradigms, which rely on sequential execution and shared
%state, struggle to capture the asynchronous nature of distributed
%systems accurately. As a result, developers may inadvertently overlook
%edge cases or race conditions, leading to unintended behaviors and
%system failures. [NON É MEGLIO EVITARE QUESTO? QUA ABBIAMO TUTTO
%SINCRONO]

Choreographic programming~\cite{M23} is an emerging programming
paradigm for distributed systems in which programs, referred to as
choreographies, specify communication protocols from a global
perspective. Instead of programming the behaviour of each component
independently or relying on a central orchestrator, choreographic
languages describe the interactions among participants directly. A
choreography therefore abstracts away from the internal implementation
of individual components and focuses on the communication structure
that governs their coordination.
%
In this approach, the primary concern is not the arbitrary
interleavings of local computations but the communication structure of
the system: which participants interact, in which order, and under
which conditions. Choreographies make these interaction patterns
explicit at the level of the program.
%
Choreographic programming naturally applies to distributed systems
whose behaviour can be described through structured interaction
protocols. Systems whose dynamics rely heavily on loosely coordinated,
highly unstructured, or unpredictable communication patterns are
generally less amenable to this approach. Within its intended scope,
however, choreographic programming has proved effective in practice,
since choreographies enable the automatic generation of decentralised
implementations that are correct by construction~\cite{CM13}.

% { In the kinds of distributed applications we can apply our
% technique on, the behaviour of the system is largely determined by
% how components communicate, rather than by arbitrary interleavings
% of independent local steps.  What matters is which participants
% interact, in which order, and under which conditions, and this is
% exactly the information that a choreography makes explicit.  Our
% setting is therefore designed towards systems whose main dynamics
% can be described in terms of such structured interaction patterns,
% while more loosely coordinated or highly unstructured behaviours are
% deliberately not addressed by the framework.  }

This paper presents a choreographic language designed for modelling
concurrent probabilistic systems and introduces a {\em projection}
function for translating protocols described in this language into
PRISM code. This choreographic approach not only simplifies the
modelling process but also ensures integration with PRISM's powerful
analysis capabilities.
%
In order to illustrate the idea concretely, consider the following
choreography:
%
\[
  \begin{array}{lll}
    C \stackrel{\mathsf{def}}{=} \interactBase{\nr}{\nrr}
    \left\{
    \begin{array}{lll}
      \lambda_1: \ (x'=1)\ \&\ (y'=2);\ C\\[1mm]
      \lambda_2: \ (x'=3)\ \&\ (y'=1);\ C
    \end{array}
    \right.
  \end{array}
\]
%
This specification recursively defines an interaction where process
\(\role{p}\) communicates with process \(\role{q}\), choosing the
first branch with rate $\lambda_1$ (and setting \(x\) to 1 and \(y\)
to 2) or the second branch with rate $\lambda_2$ (and setting \(x\) to
3 and \(y\) to 1). Here, we adopt PRISM notation where the prime
symbol, e.g., $x'$, $y'$, indicates the next-state value of the
corresponding variable, that is, the value it will take after the
transition occurs.
% 
The projection mechanism translates this global choreography into
local PRISM modules by annotating each branch with a unique label
(e.g., $a$ and $b$) and using a state variable (e.g., \(s_{\role p}\)
and \(s_{\role q}\)) to uniquely identify each step.
%
One possible representation of the behaviour described by the
choreography in PRISM is the following:
% The behaviour described by the choreography can be represented in
% PRISM as:
%
\[
\begin{array}{llll}
\role{p}: \{ & \quad \commandBase {a}{s_{\role p}=0} {\lambda_1:(x'=1)\ \&\ (s_{\role p}'=1)} \quad\qquad \commandBase {}{s_{\role p}=1} {1:(s_{\role p}'=0)}\\[2mm]
              & \quad \commandBase {b}{s_{\role p}=0} {\lambda_2:(x'=3)\ \&\ (s_{\role p}'=2)} \quad\qquad \commandBase {}{s_{\role p}=2} {1:(s_{\role p}'=0)} \quad \}\\[4mm]
\role{q}: \{ & \quad \commandBase {a}{s_{\role q}=0} {1:(y'=2)\ \&\ (s_{\role q}'=1)} \qquad\quad \commandBase {}{s_{\role q}=1} {1:(s_{\role q}'=0)}\\[2mm]
              & \quad \commandBase {b}{s_{\role q}=0} {1:(y'=1)\ \&\ (s_{\role q}'=2)} \qquad\quad \commandBase {}{s_{\role q}=2} {1:(s_{\role q}'=0)} \quad \}\\
\end{array}
\]
%
In this translation, modules $\nr$ and $\nrr$ contain a set of
commands. For example,
$\commandBase {a}{s_{\role p}=0} {\lambda_1:(x'=1)\ \&\ (s_{\role
    p}'=1)}$ in module $\nr$ is a command where $a$ is a label over
which the module can synchronise with other modules, $s_{\role p}=0$
is a guard, $\lambda_1$ is a rate, and $(x'=1)\ \&\ (s_{\role p}'=1)$
is a list of commands for modifying the state. Here, we use labels $a$
and $b$ for simulating the two branches of the choreography.
%
The command labeled \(a\) in \(\role{p}\)’s module corresponds to the
branch with rate \(\lambda_1\), and the matching command in
\(\role{q}\)’s module (with rate \(1\)) sets \(y\) to 2; similarly,
label \(b\) corresponds to the second branch with rate \(\lambda_2\)
for \(\role{p}\) and rate \(1\) for \(\role{q}\), setting \(x\) and
\(y\) appropriately.  After executing an interaction command in each
module, a subsequent reset command (with rate 1) sets the state
counters \( s_{\role p} \) and \( s_{\role q} \) back to 0 due to the
recursive call following the branching. This mechanism ensures that
these commands are only executable when the system has reached this
particular state.



Through our contributions, we aim to provide a smooth workflow for
modelling, analysing, and verifying concurrent probabilistic systems,
ultimately increasing their usability in various application domains.
%
In particular, by employing choreographies, we gain a clear and
comprehensive view of the interactions occurring within the system,
allowing us to discern the flow of processes and detect any potential
sources of error in the modelling phase.
% This transparency ensures that each step is understood and executed
% accurately, minimizing the likelihood of errors and enhancing the
% reliability of the system.


\paragraph{Contributions and Overview.} We summarise our contributions as
follows:
\begin{itemize} 
\item we propose a choreographic language equipped with well-defined
  syntax and semantics, tailored specifically for describing
  concurrent systems with probabilistic behaviours
  (\S~\ref{sec:chor}); % To the best of our knowledge, this is the first
  % probabilistic choreography language that is not a type abstraction;

\item we introduce a semantics for the minimal fragment of PRISM
  needed for code generation from choreographies (\S~\ref{sec:prism}),
  which follows PRISM's original semantics~\cite{PRISMdoc};
  % enhancing its usability and facilitating clearer understanding for
  % users.

\item we establish a rigorous definition for a translation function
  from choreographies to PRISM (\S~\ref{sec:proj}), and address its
  correctness. This translation serves as a crucial intermediary step
  in transforming models described in our choreographic language into
  PRISM-compatible representations;

\item we give an implementation~\cite{repository} that executes the
  defined translation function, enabling users to utilise PRISM's
  robust analysis features while benefiting from the expressiveness of
  choreographies;

\item we show some use cases in order to demonstrate the applicability
  of our language (\S~\ref{sec:bench}).
\end{itemize}

% \paragraph{Changes with respect to the conference version.}
% %
% This article is an expanded version of our paper presented at
% COORDINATION 2024~\cite{CV24}. It includes detailed definitions and a
% proof, which were previously omitted. Additionally, we provide a
% thorough analysis of richer language constructs incorporated in the
% implementation. Furthermore, we introduce a new set of use cases,
% analysed through choreographies, including an example that illustrates
% when the PRISM endpoint setting can be just as effective as a
% choreography-based approach.

\paragraph{Changes with respect to the conference version.}
%
This article is an expanded version of our paper presented at
COORDINATION 2024~\cite{CV24}. It includes detailed definitions and a
complete proof of the main theorem, which were previously omitted. We
also present a revised projection function that provides a cleaner
formalisation and no longer relies on counters for the control state.
Furthermore, we introduce additional language constructs, which are
also supported by the implementation. Additionally, we provide a more
extensive discussion of the language features supported by the
implementation and expand the related work section. Finally, we
introduce new use cases analysed through choreographies, including
case studies on a simple peer-to-peer protocol, synchronous leader
election, and the Dining Cryptographers protocol.

%%% Local Variables: 
%%% mode: latex
%%% TeX-master: "main"
%%% End:
